---
title: "Introduction to Sliding Window Deconvolution"
date: "July 23, 2021"
author: "Emily LaPlante"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(XDecCHI)
```

This package assumes you have already completed a deconvolution which provides estimated proportions for each cell type. You will need:

1) An expression matrix where features are rows (in the example genes) and samples are columns
2) A proportion matrix where samples are rows and cell types are columns

Example data provided are 45 ICGC samples across two genes, MSLN and RELA.

There are two functions you can perform.

1) Correlate 2 genes of interest in a cell type specific manner
2) Identify all genes that correlate to your gene of interest in a cell type specific manner

#Correlate 2 genes of interest

```{r run 2Gene 2CT}

Cor2Genes=slidingWindow_2Gene_2CT(GeneExp = SlidingWindowExtension::ExpressionMatrix,
                                  CTproportions = SlidingWindowExtension::ProportionMatrix,
                                  GOI_anchor = "MSLN", #Gene samples are ordered by
                                  GOI_cor = "RELA", #Gene you want to correlate to
                                  CT_anchor = "Classical.A", #Cell type of anchor gene
                                  CT_cor="Immune") #Cell type of correlated gene

```

## Output of 2 gene correlation {.tabset}

###Written output

The function reports the correlation and p.value of the two genes

```{r}

Cor2Genes$Correlation
Cor2Genes$p.value

```

###Cell type specific graphs

A plot with the correlation of the cell type specific expression across the windows
A plot with estimated expression across all windows for the anchor gene
A plot with estimated expression across all windows for the correlated gene

```{r}

Cor2Genes$Cor_plot
Cor2Genes$CT_anchor_expression
Cor2Genes$CT_cor_expression

```

###Bulk expression graphs

A plot with bulk expression of the anchor gene of all samples ordered as they were for the sliding window deconvolution 
A plot with bulk expression of the correlated gene of all samples ordered as they were for the sliding window deconvolution 

```{r}

Cor2Genes$Bulk_anchor_expression
Cor2Genes$Bulk_cor_expression

```

#Correlate 2 genes of interest

In this case, you have 1 gene of interest and are hoping to ID

```{r run 1 gene 2CT}

Cor2CT=slidingWindow_1Gene_CT_correlations(GeneExp = SlidingWindowExtension::ExpressionMatrix,
                                  CTproportions = SlidingWindowExtension::ProportionMatrix,
                                  GOI_anchor = "MSLN", #Gene samples are ordered by
                                  CT_anchor = "Classical.A", #Cell type of anchor gene
                                  CT_cor="Immune", #Cell type of correlated gene
                                  Type_pValue="p.value", #FDR or p.value cut off
                                  Significance_cutoff=0.05, #Significance < this to be reported
                                  Correlation_cutoff=0.8) #Correlation > this to be reported (- and +)

```

## Output of 2 cell type {.tabset}

The function reports the correlation and p.value of the two genes that pass the thresholds

###All that pass

This reports all features that pass the significance and correlation thresholds regardless of if the correlation is positive or negate

```{r}

Cor2CT$all_correlations

```

###Positive correlations only

This matrix only reports positive correlations that pass the significance and correlation thresholds

```{r}
Cor2CT$Positive_correlations
```

###Negative correlations only

This matrix only reports negative correlations that pass the significance and correlation thresholds

```{r}
Cor2CT$Negative_correlations
```

